---
title: "frisch_waugh"
output: html_document
---

```{r}
library("dplyr")
library("tidyr")
library("ggplot2")

set.seed(2)
rnorm_data = rnorm(1000, 10, 1)
plt_x = seq(min(rnorm_data), max(rnorm_data),length=500)
plt_density_curve = dnorm(plt_x, mean=mean(rnorm_data), sd=sd(rnorm_data))
hist(rnorm_data, freq=FALSE)
lines(plt_x, plt_density_curve,type="l",col="blue")
```

```{r}
n = 10000
# x1 = runif(n, min=-100, max=100)
x1 = seq(-100, 100,length=n)
x2 = rnorm(n, mean=2, sd=10) + (0.01*x1)*(0.01*x1)
x3 = as.numeric(x1 > 20 & x1 < 50)*(300 + 2*x2 + 3*x2*x2 + sqrt(abs(x1)))
y = 2*x1 + 7*x2 + x3
df = data.frame(x1, x2, x3, y)
```

```{r}
ggplot(df) + geom_point(aes(x=x1,y=y,color=x2))
```

```{r}
ggplot(df) + geom_point(aes(x=x2,y=y,color=x3))
```

Step 1 purges the effects of the regressors in X1 on the regressors in X2
Step 2 purges the effects of the regressors in X1 on Y.
Step 3 estimates the effect of the regressors in X2 on Y using the parts in X2 and
Y that have excluded the effects of X1.

• Imperfect multicollinearity does not affect the unbiasedness of the OLS estimators.
However, it does affect the efficiency, i.e., the variance of the OLS estimators
```{r}
lm(y ~ x1 + x2 + x3, data=df)
```

```{r}
output_r1 = lm(y ~ x1 + x3, data=df)
output_r1_coefs = output_r1$coefficients
output_r1_fittedvalues = output_r1$fitted.values
output_r1_residuals = output_r1$residuals
df$r1_fitted = output_r1_fittedvalues
df$r1_residuals = output_r1_residuals
output_r1
```

```{r}
output_r2 = lm(x2 ~ x1 + x3, data=df)
output_r2_residuals = output_r2$residuals
df$r2_residuals = output_r2_residuals
df$r2_fitted = output_r2$fitted.values
output_r2
```

```{r}
output_r3 = lm(r1_residuals ~ r2_residuals, data=df)
df$r3_fitted = output_r3$fitted.values
df$r3_residuals = output_r3$residuals
output_r3
```

```{r}
ggplot(df) + 
  geom_point(aes(x=x1,y=y),color="blue") +
  geom_point(aes(x=x1,y=r1_fitted),color="red") +
  geom_point(aes(x=x1,y=r1_residuals),color="green")

ggplot(df) + 
  geom_point(aes(x=x1,y=x2),color="blue") +
  geom_point(aes(x=x1,y=r2_fitted),color="red") +
  geom_point(aes(x=x1,y=r2_residuals),color="green")

ggplot(df) + 
  geom_point(aes(x=r2_residuals,y=r1_residuals),color="blue") +
  geom_point(aes(x=r2_residuals,y=r3_fitted),color="red") +
  geom_point(aes(x=r2_residuals,y=r3_residuals),color="green")

```

